---
title: "Heatmap unique to salmon"
author: "Audrey Hendricks"
date: "4/27/2022"

output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '6'
    toc_float: true
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/sakaizarasolofomananarajery/Desktop/Research/Hendriks Folder/Biotransformers/")

library(ggplot2)
library(qqplotr)
library(lme4)
library(lmerTest)
library(cowplot)
library(here)
library(readr)
library(tidyverse)
library("pheatmap")
here()
```

# lipid
```{r}
foodtype <- "salmon"
datatype <- "lipid_bio"
file.name <- paste(foodtype, datatype, sep = "_")
```



read in lipid data
```{r}
diet_lipid<-read_csv(paste("food_table/Change_in_Ranks_Primary_Analysis_",file.name,".csv",sep=""))
health_change_lipid<-read_csv(paste("food_table/change_in_Health_Metabolites_",file.name,".csv",sep=""))

diet_lipid$est_std_diet<-diet_lipid$Estimate_Intercept/diet_lipid$Std_Error_Intercept
```


reformatting data
```{r}
health_change_lipid2<-health_change_lipid[,c("Metabolite", "type", "Estimate_Metabolite", "Std_Error_Metabolite", "p_value_Metabolite", "FDR_Metabolite")]
health_change_lipid2$est_std<-health_change_lipid2$Estimate_Metabolite/health_change_lipid2$Std_Error_Metabolite


health_change_lipid_wide<- pivot_wider(health_change_lipid2, names_from = "type", values_from = c("Estimate_Metabolite", "Std_Error_Metabolite", "p_value_Metabolite", "FDR_Metabolite", "est_std"), names_sep = ".")



#health_change_lipid_wide<-reshape(health_change_lipid2, idvar="Metabolite", timevar="type", direction="wide")

total_lipid<-merge(diet_lipid[,c("Metabolite", "Estimate_Intercept", "Std_Error_Intercept", "p_value_Intercept", "FDR_Intercept","est_std_diet","Alignment.Value", "Annotations", "Compound.Name", "Formula", "Ionization.mode", "Mass","Retention.Time")], health_change_lipid_wide, by="Metabolite")

total_lipid_sig<-total_lipid[total_lipid$FDR_Intercept<0.05 | total_lipid$FDR_Metabolite.SBP<0.05 | total_lipid$FDR_Metabolite.DBP<0.05 | total_lipid$FDR_Metabolite.Insulin<0.05 | total_lipid$FDR_Metabolite.Chol_T<0.05 | total_lipid$FDR_Metabolite.TG<0.05 | total_lipid$FDR_Metabolite.LDL<0.05 | total_lipid$FDR_Metabolite.HDL<0.05 | total_lipid$FDR_Metabolite.CRP<0.05 | total_lipid$FDR_Metabolite.ApoB<0.05 | total_lipid$FDR_Metabolite.HDL_LDL<0.05 | total_lipid$FDR_Metabolite.Chol_T_HDL <0.05,]

total_lipid_sig_nodiet<-total_lipid[total_lipid$FDR_Metabolite.SBP<0.05 | total_lipid$FDR_Metabolite.DBP<0.05 | total_lipid$FDR_Metabolite.Insulin<0.05 | total_lipid$FDR_Metabolite.Chol_T<0.05 | total_lipid$FDR_Metabolite.TG<0.05 | total_lipid$FDR_Metabolite.LDL<0.05 | total_lipid$FDR_Metabolite.HDL<0.05 | total_lipid$FDR_Metabolite.CRP<0.05 | total_lipid$FDR_Metabolite.ApoB<0.05 | total_lipid$FDR_Metabolite.HDL_LDL<0.05 | total_lipid$FDR_Metabolite.Chol_T_HDL <0.05,]

total_lipid_sig_onlydiet<-total_lipid[total_lipid$FDR_Intercept<0.05,]

diet_sig<-total_lipid[total_lipid$FDR_Intercept<0.05,]

compounds_of_interest <- diet_sig[diet_sig$FDR_Metabolite.SBP<0.05 | diet_sig$FDR_Metabolite.DBP<0.05 | diet_sig$FDR_Metabolite.Insulin<0.05 | diet_sig$FDR_Metabolite.Chol_T<0.05 | diet_sig$FDR_Metabolite.TG<0.05 | diet_sig$FDR_Metabolite.LDL<0.05 | diet_sig$FDR_Metabolite.HDL<0.05 | diet_sig$FDR_Metabolite.CRP<0.05 | diet_sig$FDR_Metabolite.ApoB<0.05 | diet_sig$FDR_Metabolite.HDL_LDL<0.05 | diet_sig$FDR_Metabolite.Chol_T_HDL <0.05,]

#All 74 metabolites
total_lipid_heatmap<-total_lipid[,c("est_std_diet", "est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]

#Any of the 55 with at least significant association to one of the health outcome or diet
total_lipid_sig_heatmap<-total_lipid_sig[,c("est_std_diet", "est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]

#Only Metabolites (49) with significant change after diet
total_lipid_sig_diet_heatmap<-total_lipid_sig_onlydiet[,c("est_std_diet", "est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]

#Any metabolite(16) that have at least one significant association to health outcomes, do not plot diet.
total_lipid_sig_nodiet_heatmap<-total_lipid_sig_nodiet[,c("est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]

#Any metabolite (16) that have at least one significant association to health outcomes, plot diet.
total_lipid_sig_alsodiet_heatmap<-total_lipid_sig_nodiet[,c("est_std_diet", "est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]

#Metabolites whose change are significant and significantly associated with health outcomes.
compounds_of_interest_heatmap <- compounds_of_interest[,c("est_std_diet", "est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]

colnames(total_lipid_sig_alsodiet_heatmap)<-colnames(total_lipid_heatmap)<-colnames(total_lipid_sig_heatmap)<-colnames(total_lipid_sig_diet_heatmap)<-c("Diet", "SBP","DBP", "Insulin", "Chol_T", "TG", "LDL", "HDL", "CRP", "ApoB", "HDL / LDL", "Chol_T / HDL")

colnames(total_lipid_sig_nodiet_heatmap)<-c("SBP","DBP", "Insulin", "Chol_T", "TG", "LDL", "HDL", "CRP", "ApoB", "HDL / LDL", "Chol_T / HDL")

colnames(compounds_of_interest_heatmap)<- c("Diet", "SBP","DBP", "Insulin", "Chol_T", "TG", "LDL", "HDL", "CRP", "ApoB", "HDL / LDL", "Chol_T / HDL")

row.names(compounds_of_interest_heatmap) <- compounds_of_interest$Compound.Name
row.names(total_lipid_sig_heatmap)<-total_lipid_sig$Compound.Name
row.names(total_lipid_sig_diet_heatmap)<-total_lipid_sig_onlydiet$Compound.Name
row.names(total_lipid_sig_nodiet_heatmap)<-row.names(total_lipid_sig_alsodiet_heatmap)<-total_lipid_sig_nodiet$Compound.Name
row.names(total_lipid_heatmap)<-total_lipid$Compound.Name

```

## heatmap

### Any of the 55 with at least a significant association to one of the health outcome or diet with names
```{r}
#set similar scale for all heatmap
rg <- max(abs(as.matrix(compounds_of_interest_heatmap)))

#Any of the 55 with at least a significant association to one of the health outcome or diet
sig_heatmap <- pheatmap(as.matrix(total_lipid_sig_heatmap),breaks = seq(-rg, rg, length.out = 100), cutree_rows = 3, cutree_cols  = 2, fontsize_row = 6,fontsize_col = 12, show_rownames=F) 
ggsave2(filename=paste0("sig_heatmap_", file.name,".jpeg", sep=""), plot=sig_heatmap, device="jpeg",
          path=paste0("./heatmap"))
```

### Any of the 55 with at least a significant association to one of the health outcome or diet with names
```{R}
#Any of the 55 with at least a significant association to one of the health outcome or diet with names
sig_heatmap_withnames <- pheatmap(as.matrix(total_lipid_sig_heatmap),breaks = seq(-rg, rg, length.out = 100), cutree_rows = 3, cutree_cols  = 2, fontsize_row = 6,fontsize_col = 12, show_rownames=T) 
ggsave2(filename=paste0("sig_heatmap_named_", file.name,".jpeg", sep=""), plot=sig_heatmap_withnames, device="jpeg",
          path=paste0("./heatmap"))
```


### Any metabolite(16) that have at least one significant association to health outcomes, do not plot diet
```{r}
#Any metabolite(16) that have at least one significant association to health outcomes, do not plot diet
sig_nodiet_heatmap<- pheatmap(as.matrix(total_lipid_sig_nodiet_heatmap),breaks = seq(-rg, rg, length.out = 100), cutree_rows = 2, cutree_cols  = 2, fontsize_row = 6,fontsize_col = 12, show_rownames=T)
ggsave2(filename=paste0("sig_nodiet_heatmap_", file.name,".jpeg", sep=""), plot=sig_nodiet_heatmap, device="jpeg",
          path=paste0("./heatmap"))
```


### Only Metabolites (49) with significant change after diet
```{R}
#Only Metabolites (49) with significant change after diet
# pheatmap(as.matrix(total_lipid_sig_diet_heatmap), cutree_rows = 2, cutree_cols  = 3, fontsize = 12, show_rownames=F) 
sig_diet_heatmap<- pheatmap(as.matrix(total_lipid_sig_diet_heatmap), breaks = seq(-rg, rg, length.out = 100), cutree_rows = 2, cutree_cols  = 3, fontsize = 12, show_rownames=F) 
```

### Only Metabolites (49) with significant change after diet with names
```{r}
#Only Metabolites (49) with significant change after diet with names
sig_diet_heatmap_withnames<- pheatmap(as.matrix(total_lipid_sig_diet_heatmap), breaks = seq(-rg, rg, length.out = 100), cutree_rows = 2, cutree_cols  = 3, fontsize = 6, show_rownames=T) 
ggsave2(filename=paste0("sig_diet_heatmap_", file.name,".jpeg", sep=""), plot=sig_diet_heatmap, device="jpeg",
          path=paste0("./heatmap"))

ggsave2(filename=paste0("sig_diet_heatmap_named_", file.name,".jpeg", sep=""), plot=sig_diet_heatmap_withnames, device="jpeg",
          path=paste0("./heatmap"))
```


### Any metabolite (16) that have at least one significant association to health outcomes, plot diet 
```{r}
#Any metabolite (16) that have at least one significant association to health outcomes, plot diet
sig_alsodiet_heatmap<- pheatmap(as.matrix(total_lipid_sig_alsodiet_heatmap),breaks = seq(-rg, rg, length.out = 100), cutree_rows = 2, cutree_cols  = 3, fontsize_row = 6,fontsize_col = 12, show_rownames=T)
ggsave2(filename=paste0("sig_alsodiet_heatmap_", file.name,".jpeg", sep=""), plot=sig_alsodiet_heatmap, device="jpeg",
          path=paste0("./heatmap"))

```


### Metabolites whose change are significant and significantly associated with health outcomes. 
```{R}
#Metabolites whose change are significant and significantly associated with health outcomes.
cmpd_intrst_heatmap<- pheatmap(as.matrix(compounds_of_interest_heatmap),breaks = seq(-rg, rg, length.out = 100), cutree_rows = 2, cutree_cols  = 3, fontsize_row = 6,fontsize_col = 12, show_rownames=T)
ggsave2(filename=paste0("cmpd_intrst_heatmap_", file.name,".jpeg", sep=""), plot=cmpd_intrst_heatmap, device="jpeg",
          path=paste0("./heatmap"))
```

# Aqueous
```{r}
foodtype <- "salmon"
datatype <- "aqueous_bio"
file.name <- paste(foodtype, datatype, sep = "_")
```

read in aqueous data
```{r}
diet_aq<-read_csv(paste("food_table/Change_in_Ranks_Primary_Analysis_",file.name,".csv",sep=""))
health_change_aq<-read_csv(paste("food_table/change_in_Health_Metabolites_",file.name,".csv",sep=""))

diet_aq$est_std_diet<-diet_aq$Estimate_Intercept/diet_aq$Std_Error_Intercept
```

reformatting data
```{r}
health_change_aq2<-health_change_aq[,c("Metabolite", "type", "Estimate_Metabolite", "Std_Error_Metabolite", "p_value_Metabolite", "FDR_Metabolite")]
health_change_aq2$est_std<-health_change_aq2$Estimate_Metabolite/health_change_aq2$Std_Error_Metabolite


health_change_aq_wide<- pivot_wider(health_change_aq2, names_from = "type", values_from = c("Estimate_Metabolite", "Std_Error_Metabolite", "p_value_Metabolite", "FDR_Metabolite", "est_std"), names_sep = ".")

total_aq<-merge(diet_aq[,c("Metabolite", "Estimate_Intercept", "Std_Error_Intercept", "p_value_Intercept", "FDR_Intercept","est_std_diet","Alignment.Value", "Annotations", "Compound.Name", "Formula", "Ion.Species", "Mass",                 "Retention.Time")], health_change_aq_wide, by="Metabolite")

total_aq_sig<-total_aq[total_aq$FDR_Intercept<0.05 | total_aq$FDR_Metabolite.SBP<0.05 | total_aq$FDR_Metabolite.DBP<0.05 | total_aq$FDR_Metabolite.Insulin<0.05 | total_aq$FDR_Metabolite.Chol_T<0.05 | total_aq$FDR_Metabolite.TG<0.05 | total_aq$FDR_Metabolite.LDL<0.05 | total_aq$FDR_Metabolite.HDL<0.05 | total_aq$FDR_Metabolite.CRP<0.05 | total_aq$FDR_Metabolite.ApoB<0.05 | total_aq$FDR_Metabolite.HDL_LDL<0.05 | total_aq$FDR_Metabolite.Chol_T_HDL <0.05,]

total_aq_sig_nodiet<-total_aq[total_aq$FDR_Metabolite.SBP<0.05 | total_aq$FDR_Metabolite.DBP<0.05 | total_aq$FDR_Metabolite.Insulin<0.05 | total_aq$FDR_Metabolite.Chol_T<0.05 | total_aq$FDR_Metabolite.TG<0.05 | total_aq$FDR_Metabolite.LDL<0.05 | total_aq$FDR_Metabolite.HDL<0.05 | total_aq$FDR_Metabolite.CRP<0.05 | total_aq$FDR_Metabolite.ApoB<0.05 | total_aq$FDR_Metabolite.HDL_LDL<0.05 | total_aq$FDR_Metabolite.Chol_T_HDL <0.05,]

total_aq_sig_onlydiet<-total_aq[total_aq$FDR_Intercept<0.05,]

total_aq_heatmap<-total_aq[,c("est_std_diet", "est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]
total_aq_sig_heatmap<-total_aq_sig[,c("est_std_diet", "est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]
total_aq_sig_diet_heatmap<-total_aq_sig_onlydiet[,c("est_std_diet", "est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]
total_aq_sig_nodiet_heatmap<-total_aq_sig_nodiet[,c("est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]
total_aq_sig_alsodiet_heatmap<-total_aq_sig_nodiet[,c("est_std_diet", "est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]

colnames(total_aq_sig_alsodiet_heatmap)<-colnames(total_aq_heatmap)<-colnames(total_aq_sig_heatmap)<-colnames(total_aq_sig_diet_heatmap)<-c("Diet", "SBP","DBP", "Insulin", "Chol_T", "TG", "LDL", "HDL", "CRP", "ApoB", "HDL / LDL", "Chol_T / HDL")

colnames(total_aq_sig_nodiet_heatmap)<-c("SBP","DBP", "Insulin", "Chol_T", "TG", "LDL", "HDL", "CRP", "ApoB", "HDL / LDL", "Chol_T / HDL")

row.names(total_aq_sig_heatmap)<-total_aq_sig$Compound.Name
row.names(total_aq_sig_diet_heatmap)<-total_aq_sig_onlydiet$Compound.Name
row.names(total_aq_sig_nodiet_heatmap)<-row.names(total_aq_sig_alsodiet_heatmap)<-total_aq_sig_nodiet$Compound.Name
row.names(total_aq_heatmap)<-total_aq$Compound.Name
```

## heatmap

### 11 metabolites with significant change over diet intervention
```{r}
#rg <- max(abs(as.matrix(total_aq_sig_heatmap)))

# 11 metabolites with significant change over diet intervention
sig_heatmap <- pheatmap(as.matrix(total_aq_sig_heatmap),breaks = seq(-rg, rg, length.out = 100), cutree_rows = 3, cutree_cols  = 2, fontsize_row = 6,fontsize_col = 12, show_rownames=T) 
ggsave2(filename=paste0("sig_heatmap_", file.name,".jpeg", sep=""), plot=sig_heatmap, device="jpeg",
          path=paste0("./heatmap"))
```

