---
title: "Heatmap unique to salmon"
author: "Audrey Hendricks"
date: "4/27/2022"

output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '6'
    toc_float: true
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(qqplotr)
library(lme4)
library(lmerTest)
library(cowplot)
library(here)
here()
```

# lipid
read in data
```{r}
# Importing data set
diet_lipid<-read.csv("Lipid_Change_diet_Ranks_Primary_Analysis.csv")
health_change_lipid<-read.csv("Lipid_Change_in_Health_Metabolites.csv")

#creating intercept variable to use for heatmap 
diet_lipid$est_std_diet<-diet_lipid$Estimate_Intercept/diet_lipid$Std_Error_Intercept
```


reformatting data
```{r}
#Selecting column needed
health_change_lipid2<-health_change_lipid[,c("Metabolite", "type", "Estimate_Metabolite", "Std_Error_Metabolite", "p_value_Metabolite", "FDR_Metabolite")]

#Creating Metabolite variable to use for heatmap
health_change_lipid2$est_std<-health_change_lipid2$Estimate_Metabolite/health_change_lipid2$Std_Error_Metabolite


health_change_lipid_wide<-reshape(health_change_lipid2, idvar="Metabolite", timevar="type", direction="wide")

#Merging diet and health change data together
total_lipid<-merge(diet_lipid[,c("Metabolite","Estimate_Intercept","Std_Error_Intercept","p_value_Intercept","FDR_Intercept","Compound.Name","CASNumber","Number.Passed","Retention.Time","Measured.Retention.Time","Product.Mass","QualifierMass","est_std_diet")], health_change_lipid_wide, by="Metabolite")

#All (59) metabolites with either significant change after diet intervention or significant association to any health outcome  at FDR < 0.05
total_lipid_sig<-total_lipid[total_lipid$FDR_Intercept<0.05 | total_lipid$FDR_Metabolite.SBP<0.05 | total_lipid$FDR_Metabolite.DBP<0.05 | total_lipid$FDR_Metabolite.Insulin<0.05 | total_lipid$FDR_Metabolite.Chol_T<0.05 | total_lipid$FDR_Metabolite.TG<0.05 | total_lipid$FDR_Metabolite.LDL<0.05 | total_lipid$FDR_Metabolite.HDL<0.05 | total_lipid$FDR_Metabolite.CRP<0.05 | total_lipid$FDR_Metabolite.ApoB<0.05 | total_lipid$FDR_Metabolite.HDL_LDL<0.05 | total_lipid$FDR_Metabolite.Chol_T_HDL <0.05,]

# All metabolites (9) with significant association to a health outcome at FDR < 0.05
total_lipid_sig_nodiet<-total_lipid[total_lipid$FDR_Metabolite.SBP<0.05 | total_lipid$FDR_Metabolite.DBP<0.05 | total_lipid$FDR_Metabolite.Insulin<0.05 | total_lipid$FDR_Metabolite.Chol_T<0.05 | total_lipid$FDR_Metabolite.TG<0.05 | total_lipid$FDR_Metabolite.LDL<0.05 | total_lipid$FDR_Metabolite.HDL<0.05 | total_lipid$FDR_Metabolite.CRP<0.05 | total_lipid$FDR_Metabolite.ApoB<0.05 | total_lipid$FDR_Metabolite.HDL_LDL<0.05 | total_lipid$FDR_Metabolite.Chol_T_HDL <0.05,]

#all (55) metabolites that significantly change after diet intervention at FDR<0.05
total_lipid_sig_onlydiet<-total_lipid[total_lipid$FDR_Intercept<0.05,]

#All 104 lipid plotted with diet and health association
total_lipid_heatmap<-total_lipid[,c("est_std_diet","est_std.SBP","est_std.DBP","est_std.Insulin","est_std.Chol_T","est_std.TG","est_std.LDL","est_std.HDL","est_std.CRP","est_std.ApoB","est_std.HDL_LDL","est_std.Chol_T_HDL")]

#All (59) metabolites with either significant change after diet intervention or significant association to any health outcome  at FDR < 0.05
total_lipid_sig_heatmap<-total_lipid_sig[,c("est_std_diet","est_std.SBP","est_std.DBP","est_std.Insulin","est_std.Chol_T","est_std.TG","est_std.LDL","est_std.HDL","est_std.CRP","est_std.ApoB","est_std.HDL_LDL","est_std.Chol_T_HDL")]

#all (55) metabolites that significantly change after diet intervention at FDR<0.05
total_lipid_sig_diet_heatmap<-total_lipid_sig_onlydiet[,c("est_std_diet", "est_std.SBP", "est_std.DBP", "est_std.Insulin", "est_std.Chol_T", "est_std.TG", "est_std.LDL", "est_std.HDL", "est_std.CRP", "est_std.ApoB", "est_std.HDL_LDL", "est_std.Chol_T_HDL")]

# All metabolites (9) with significant association to a health outcome at FDR < 0.05, do not plot Diet
total_lipid_sig_nodiet_heatmap<-total_lipid_sig_nodiet[,c("est_std.SBP","est_std.DBP","est_std.Insulin","est_std.Chol_T","est_std.TG","est_std.LDL","est_std.HDL","est_std.CRP","est_std.ApoB","est_std.HDL_LDL","est_std.Chol_T_HDL")]

# All metabolites (9) with significant association to a health outcome at FDR < 0.05, plot diet
total_lipid_sig_alsodiet_heatmap<-total_lipid_sig_nodiet[,c("est_std_diet","est_std.SBP","est_std.DBP","est_std.Insulin","est_std.Chol_T","est_std.TG","est_std.LDL","est_std.HDL","est_std.CRP","est_std.ApoB","est_std.HDL_LDL","est_std.Chol_T_HDL")]

colnames(total_lipid_sig_alsodiet_heatmap)<-colnames(total_lipid_heatmap)<-colnames(total_lipid_sig_heatmap)<-colnames(total_lipid_sig_diet_heatmap)<-c("Diet", "SBP","DBP", "Insulin", "Chol_T", "TG", "LDL", "HDL", "CRP", "ApoB", "HDL / LDL", "Chol_T / HDL")

colnames(total_lipid_sig_nodiet_heatmap)<-c("SBP","DBP", "Insulin", "Chol_T", "TG", "LDL", "HDL", "CRP", "ApoB", "HDL / LDL", "Chol_T / HDL")

row.names(total_lipid_sig_heatmap)<-total_lipid_sig$Compound.Name
row.names(total_lipid_sig_diet_heatmap)<-total_lipid_sig_onlydiet$Compound.Name
row.names(total_lipid_sig_nodiet_heatmap)<-row.names(total_lipid_sig_alsodiet_heatmap)<-total_lipid_sig_nodiet$Compound.Name
row.names(total_lipid_heatmap)<-total_lipid$Compound.Name
```

## heatmap
PGP(18:0/22:4(72.102.132.162)) is unique M.103
```{r}
library("pheatmap")
#setting similar scale for all heatmaps
rg <- 5
```


### All (59) metabolites with either significant change after diet intervention or significant association to any health outcome  at FDR < 0.05
```{R}
pheatmap(as.matrix(total_lipid_sig_heatmap),breaks = seq(-rg, rg, length.out = 100), cutree_rows = 3, cutree_cols  = 2, fontsize_row = 6,fontsize_col = 12, show_rownames=F) 
```

### All metabolites (9) with significant association to a health outcome at FDR < 0.05, do not plot Diet
```{r}
pheatmap(as.matrix(total_lipid_sig_nodiet_heatmap),breaks = seq(-rg, rg, length.out = 100), cutree_rows = 2, cutree_cols  = 2, fontsize_row = 6,fontsize_col = 12, show_rownames=T) 
```

### All (55) metabolites that significantly change after diet intervention at FDR<0.05
```{r}
pheatmap(as.matrix(total_lipid_sig_diet_heatmap),breaks = seq(-rg, rg, length.out = 100), cutree_rows = 2, cutree_cols  = 3, fontsize = 12, show_rownames=F) 
```

### All metabolites (9) with significant association to a health outcome at FDR < 0.05, plot Diet
```{R}
pheatmap(as.matrix(total_lipid_sig_alsodiet_heatmap),breaks = seq(-rg, rg, length.out = 100), cutree_rows = 2, cutree_cols  = 3, fontsize_row = 6,fontsize_col = 12, show_rownames=T) 
```

